\documentclass[aps,prb,twocolumn,superscriptaddress,nobibnotes]{revtex4-2} % APS journal style
\usepackage[a4paper,margin=0.8in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath, amssymb}
\usepackage{hyperref}
\usepackage{float}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{titlesec} 
\usepackage[utf8]{inputenc}
\usepackage{enumitem}
\raggedbottom
% Bibliography commands
\bibliographystyle{apsrev4-2}
\titleformat{\section}{\bfseries\large\uppercase}{\thesection.}{1em}{}
\titleformat{\subsection}{\bfseries}{\thesubsection}{1em}{}
\begin{document}

\title{QEC strategies against Atom Loss in Neutral Atom Arrays}
\author{Adeniyi Osinowo}
\affiliation{
 Department of Physics and Astronomy,
 University College London,
 Gower Street,
 London,
 WC1E 6BT,
 UK
} 
\date{\today}
\begin{abstract}
% Write abstract based on progress report content
% Quantum error correction (QEC) is a critical component in the development of reliable quantum computing systems.
% Neutral atom arrays have emerged as a promising platform for quantum computing due to their scalability and long coherence times.
% However, atom loss remains a significant challenge that can compromise the effectiveness of QEC strategies.
% This progress report outlines the current state of research on QEC strategies designed to mitigate atom loss in neutral atom arrays.
% Key factors influencing atom loss, relevant error models, and potential QEC strategies are discussed.
% Preliminary results indicate that certain QEC codes show promise in addressing atom loss, but further investigation is needed to optimize their performance.
% Future work will focus on refining error models and implementing QEC strategies to enhance resilience against atom loss.
\end{abstract}
\maketitle

\section{Introduction} % Explain motivation/significance of the topic
% Paragraph: Highlight significance of topic
% Sentence: refer to development timeline to highlight rapid advancements in the field
Since their experimental realization in the early 2000s, neutral atom arrays were initially viewed as scalable but slow and low-fidelity platforms, suitable mainly for quantum simulation rather than quantum computing.
Through rapid progress in recent years, they have achieved significant milestones, including the demonstration of high-fidelity single-qubit gates, two-qubit entangling gates, and the creation of large-scale arrays with hundreds of atoms.
% Sentence: Highlight unique advantages of Neutral atom arrays over other qubit architectures
These advancements have made them a leading candidate for scalable quantum computing, offering several advantages over other qubit architectures, including scalability, long coherence times, and the ability to implement high-fidelity quantum gates.
% Sentence: Highlight challenges of atom loss in neutral atom arrays and the importance of investigating QEC strategies to mitigate the issue
However, atom loss remains a significant challenge that can compromise the effectiveness of QEC strategies on this platform.
\vspace{8pt}

% Paragraph: Define problem statement
% Sentence: Introduce neutral atom architecure to explain what of atom loss is and what causes it
Neutral atom arrays consist of individual atoms trapped in optical tweezers or optical lattices, which serve as qubits for quantum computations.
Each atom is manipulated using laser pulses to perform quantum gates, and the state of the qubits can be read out using fluorescence imaging.
During these operations, atoms can be lost due to various factors, including collisions with background gas particles, spontaneous emission, and technical imperfections in trapping mechanisms~\cite{Bluvstein2024}.
% Sentence: Explain the effects on QEC(erasure aren't like typical errors))
This leads to qubit erasure errors, which differ from standard bit-flip or phase-flip errors in that the location of the error is known, but the state of the qubit is lost.
This type of error can significantly degrade the performance of QEC codes, as it can lead to a loss of logical qubits and reduce the overall fidelity of quantum computations.
% Sentence: Explain why QEC strategies are essential for mitigating the effects of atom loss and ensuring reliable quantum computations in neutral atom arrays 
Effective QEC strategies are essential for maintaining the integrity of quantum information and ensuring reliable quantum computations in the presence of atom loss.

\vspace{8pt}
% Paragraph: Explain key research questions and how they will be explored and applied in this literature review: conduct exploratory analysis of QEC strategies that can mitigate atom loss in neutral atom arrays
%   ~ Identify key factors/challenges of atom loss in neutral atom arrays on QEC performance
%   - Identify relevant error models for atom loss in neutral atom arrays
%   - Identify relevant QEC strategies that can mitigate atom loss in neutral atom arrays
%   ~ Identify key factors to consider when assessing QEC strategies for atom loss in neutral atom arrays
This literature review aims to conduct an exploratory analysis of state-of-the-art QEC strategies that effectively mitigate qubit loss errors. 
The key research questions to be addressed include:
\begin{itemize}
    % \item What are the key challenges associated with atom loss in neutral atom arrays, and how do they impact QEC performance?
    \item What error models effectively represent atom loss?
    \item What QEC strategies have been proposed to mitigate erasure errors?
    % \item What are the key factors to consider when assessing the effectiveness of QEC strategies for atom loss in neutral atom arrays?
\end{itemize}

\section{Methodology}
% Paragraph: Explain scope and approach for reviewing literature and the sources that have been identified so far. 
This review focuses on recent research articles, review papers, and experimental studies on QEC strategies addressing erasure errors across various platforms, with the goal of identifying strategies adaptable to neutral atom arrays.
The following sections address the research questions outlined in the introduction.

% \subsection{Challenges of atom loss in neutral atom arrays on QEC performance}
% Paragraph: Discuss key factors/challenges of atom loss in neutral atom arrays on QEC performance
% Sentence: Explain how atom loss occurs in neutral atom arrays and the mechanisms behind it (e.g., collisions, spontaneous emission, technical imperfections)
% Sentence: Explain why erasure errors from atom loss are different from standard Pauli errors and how they can impact the performance of QEC codes
% Sentence: Explain the why traditional QEC codes may not be effective against erasure errors and the need for specialized QEC strategies to address this issue

\subsection{Modelling atom loss in neutral atom arrays}
% Paragraph: Discuss relevant error models for atom loss in neutral atom arrays
Studies on QEC in neutral atom arrays typically employ a combination of an atom loss model which accounts for losses during gate operations due to Rydberg excitation and depolarizing noise model, which characterizes imperfections arising from pulse applications and environmental interactions~\cite{Perrin_2025}.
The atom loss model is represented as a stochastic, gate-triggered erasure process occurring primarily during two-qubit CZ gates where each CZ gate carries a fixed probability $p_{loss}$ that either participating atom is lost. 
Once an atom is lost, all subsequent gates acting on it are removed from the circuit until it is reloaded.
The depolarizing noise model is applied to all gates, with a fixed probability $p_{depol}$ that the gate operation fails and results in a completely mixed state.
These models capture the key mechanisms of atom loss and gate imperfections in neutral atom arrays, allowing for the assessment of QEC strategies under realistic conditions.
%  - Erasure error models
%  - Stochastic loss models
%  - Loss models based on physical mechanisms of atom loss (e.g., collision-induced loss, spontaneous emission-induced loss)
%  - Hybrid models combining multiple loss mechanisms
%  - Consideration of temporal and spatial correlations in atom loss

\subsection{Atom loss mitigating QEC strategies}
% Paragraph: Discuss relevant QEC strategies that can mitigate atom loss in neutral atom arrays
%   - Loss Detection Units(LDU) & Teleportation-based LDU
%   - Decoder: Quantum Low-Density Parity-Check (QLDPC) codes with Belief Propagation and OSD Decoders
%   - Decoder: Delayed erasure decoding after loss detection
%   - Dual-rail encoding for loss detection and correction
Previous research have proposed several QEC strategies to mitigate atom loss in neutral atom arrays, including:

\begin{itemize}
    \item Teleportation-based Loss Detection Units (LDU): This approach involves using quantum teleportation to detect and correct for atom loss errors. By teleporting the state of a qubit to a new location, the system can identify if an atom has been lost and take corrective action accordingly~\cite{Perrin_2025}.
    \item Quantum Low-Density Parity-Check (QLDPC) Codes: These codes are designed to handle erasure errors effectively. They can be decoded using belief propagation and ordered statistics decoding (OSD) algorithms, which can improve the performance of QEC in the presence of atom loss~\cite{Roffe_2020}.
    \item Delayed Erasure Decoding: This strategy involves delaying the decoding process until after the loss has been detected, allowing for more accurate error correction based on the known location of the erasure~\cite{ycwc-3myc}.
    \item Dual-rail Encoding: This method encodes quantum information across two physical qubits, allowing for the detection and correction of atom loss by comparing the states of the two qubits~\cite{violaris2026developmentssuperconductingerasurequbits}.
\end{itemize}

\section{Results \& Discussion} % presentation of any preliminary results
% Paragraph: Describe what is being done so far in terms of implementation and testing of identified QEC strategies for atom loss in neutral atom arrays

Preliminary work involves implementing simple QEC codes with basic error models to develop familiarity with the Stim simulation framework~\cite{Gidney_2021} while building toward reproducing results from existing research.
A repetition code has been successfully implemented with simplified error models, confirming expected code behavior and providing a foundation for understanding QEC performance metrics. 
Current efforts focus on gradually incorporating more realistic error models from~\cite{Perrin_2025} and extending implementations to include surface codes, enabling a systematic assessment of QEC strategies under atom loss conditions.

\section{Conclusion}
% 	- Summary of progress 
This preliminary study has laid the groundwork for a comprehensive analysis of QEC strategies designed to mitigate atom loss in neutral atom arrays.
Future work will focus on reproducing and extending existing models, as well as exploring additional QEC strategies to enhance the resilience of neutral atom arrays against atom loss.
The biggest challenges encountered so far include learning to use the detector formalism in Stim and implementing more complex QEC codes, which require careful consideration of error models and decoding strategies.
With continued effort, this challenge can be overcome, and the insights gained from this research will contribute to the development of more robust quantum computing systems based on neutral atom arrays.
% 	- Outline and explain difficulties/limitations encountered so far and questions that remain unanswered
% 	- Briefly describe a revised Timeline or changes to plan

\section*{REFERENCES}
\makeatletter
\let\pre@bibdata\@empty
\makeatother
\bibliography{reference_texts} 
\end{document}